Improved Bounds for Online Balanced Repartitioning

### 2020-04-24 paper/disc20clustering.tex
Comments
did some of the comments on k=2 proof

We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
%More precisely,we charge them only the cost of the collocation, per such event.
%Hence,we charge the cost of  this collocation to OPT for the first time.
Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
which  costs $2\alpha$.
%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
paying the price of only one migration per pair.
This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
even when OPT actually pays more that that.
E.g.,
OPT can collocate two costly pairs with a single swap if it has both pairs on the cut of same two clusters.
Later, Alg collocates those pairs with two swaps (see Figure \ref{fig:TBD}).
Hence,
whenever OPT collocates a pair,
we charge OPT only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ for Alg).

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	\mahmoud{still proof sketch}
	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$ immediately after arrival of $\sigma_t$.
	Assume on request $\sigma_{t}=\set{u,v}$,
	Alg collocates the pair,
	which means $ C^t_{\set{u,v}} = \lambda$ immediately before Alg resets the counter.
	Let $\sigma_{t''}, t''<t$ be the last request (before $\sigma_{t}$)
	on which Alg collocates $\set{u,v}$.
	Let $t'$ s.t.~$t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
	in order to collocate some other pair that shares a node with $\set{u,v}$,
	denoted by $\set{u,v'}, v' \neq v$.
	%e.g., the request $\sigma_{t'} = \set{u,v'}$.
	%By Alg's definition,
	%$C^{t'}_{\set{u,v'}}  = C^t_{\set{u,v}} = \lambda$.
	We denote the set of all requests to $\set{u,v}$
	between times $t'$ and $t$ by $\sigma_{\set{u,v}}[t',t]$.
	Thus,
	$| \sigma_{\set{u,v}}[t',t] | = \lambda$.
	The overall cost of requests that increase the counter,
	i.e.~requests in $\sigma_{\set{u,v}}[t',t]$,
	and the moving cost,
	incurred by Alg is $\lambda + 2\alpha$.
  %That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.

@@

  \begin{enumerate}
  \item Case i)
	OPT has $\set{u,v}$ separated (in OPT's configuration) during $[t',t]$
	and therefore it pays $\lambda$.
	We charge OPT $\lambda/2 = \alpha$.
	The remaining cost will be used in a latter case.

  \item Case ii)  
	OPT collocates $\set{u,v}$ some time during $[t'',t]$.
	Any cost that we charge to OPT upon $\sigma_{t''}$,
	accounts for OPT's actions before $t''$.
	Therefore in this case,
	we charge OPT the (unaccounted) cost $\alpha$.

  \item Case iii)
	OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
	Obviously in this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$ and
	we can't charge it directly.
	However,
	OPT must pay $\lambda$ for all requests in $\sigma_{\set{u,v'}}[t'',t']$,
	and we charge it partially to OPT at $t'$ (as in Case i).
	Since we always charge OPT no more than $\alpha$,
	we can charge OPT another $\alpha$,
	the remaining cost of requests $\sigma_{\set{u,v'}}[t'',t']$.
	Intuitively,
	at $t$,
  we charge OPT $\alpha$ for separating $\set{u,v}$ at $t'$.
  \end{enumerate}

	It can be easily verified that any deviation from these cases only increases the cost to OPT.
	Thus,
	OPT pays at least $\alpha$ for each collocation performed by Alg,
	which gives the competitive ratio 
	$(\lambda + 2\alpha)/ \alpha = 4$.

	It remains to charge both Alg and OPT for requests that do not lead to collocations when the sequence ends.
	Consider  pairs $\set{u,v}$ and $\set{u,v'}$ and
	times $t'' < t' < t = t_m$ as defined previously.
	Assume $\sigma_{\set{u,v}}[t'', t_m] \leq \lambda - 1$.
	This means Alg has the pair separated at $t_m$
	and paid the (communication) cost $r_{\set{u,v}} := | \sigma_{\set{u,v}}[t'', t_m]|$.
	If OPT has the pair separated during $[t'',t_m]$ then it pays $r_{\set{u,v}}$.
	If OPT collocates the pair some time during $[t'', t]$ then analogous to Case (ii),
	we charge it $\alpha$.
	Otherwise,
	if OPT collocates the pair before $t''$,
	then we charge it $\alpha$ analogously to Case (iii).
	Hence, 
	$Cost(OPT)_{\set{u,v}} / Cost(Alg)_{\set{u,v}} = r_{\set{u,v}}/ \alpha \leq (\lambda-1) / \alpha < 2$.
	The claim follows from here.
\end{proof}

\maciek{$\sigma_t$ undefined (I know what it is, though).}
\maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist?}
\maciek{Define $C^t$ to use once}
\maciek{"It can be easily verified"... True, but avoid such formulation. Something along these lines: if opt migrates earlier than t', then additionally pays for X, if later than Y.}
\maciek{"Thus, OPT pays ... which gives the competitive ratio" It is too early to give the competitive ratio if in the next paragraph we consider yet another case. }
\maciek{Reason that the cost sharing between cases C1 and C3 is possible. Unclear to me, why every time C3 happens there must be a corresponding C1.}

>>>>

We charge both OPT and Alg whenever Alg collocates a pair (i.e., on a collocation event).
%More precisely,we charge them only the cost of the collocation, per such event.
%Hence,we charge the cost of  this collocation to OPT for the first time.
Alg collocates a pair always with a swap (i.e., two simultaneous migrations),
which  costs $2\alpha$.
%We charge OPT only for requests up to the one that triggers a collocation (i.e., not for later requests on the same pair).
OPT may save some costs by collocating multiple pairs at once (before they incur too much cost),
paying the price of only one migration per pair.
This observation suggests that we can't charge OPT more than $\alpha$ per collocation,
even when OPT actually pays more that that.
E.g.,
OPT can collocate two costly pairs with a single swap if
it has both pairs on the cut of  the same two clusters.
Later, Alg collocates those pairs with two swaps (see Figure \ref{fig:TBD}).
Hence,
whenever OPT collocates a pair,
we charge OPT only the cost of moving a single node to the cluster of the other,
i.e., $\alpha$ (in contrast to $2\alpha$ for Alg).

\begin{proof}[Proof of Theorem \ref{thm:k=2}]
	\mahmoud{still proof sketch}
	Let $C^{t}_{\set{u,v}}$ denote the  value of $C_{\set{u,v}}$
	immediately after arrival of the $t$th request,
	$\sigma_t$.
	Assume on request $\sigma_{t}=\set{u,v}$, 
	i.e., at (time) $t \geq 1$,
	Alg collocates the pair,
	which means $ C^t_{\set{u,v}} = \lambda$ immediately before Alg resets the counter.
	Let $\sigma_{t''}, t''<t$ be the last request (before $\sigma_{t}$)
	upon which Alg collocates $\set{u,v}$.
	Let $t'$ s.t.~$t'' < t' < t$ be the  time when Alg separates $\set{u,v}$
	in order to collocate some other pair that shares a node with $\set{u,v}$,
	denoted by $\set{u,v'}, v' \neq v$.
	In case the first collocation of $\set{u,v}$ occurs at $t$,
	if the pair is collocated in the initial configuration then
	we set $t''=0$,
	else,
	we set $ t'' = -\infty, t'=0$.
	%e.g., the request $\sigma_{t'} = \set{u,v'}$.
	%By Alg's definition,
	%$C^{t'}_{\set{u,v'}}  = C^t_{\set{u,v}} = \lambda$.
	We denote the set of all requests to $\set{u,v}$
	between times $t'$ and $t$ by $\sigma_{\set{u,v}}[t',t]$.
	Thus,
	$| \sigma_{\set{u,v}}[t',t] | = \lambda$.
	The overall cost of requests that increase the counter,
	i.e.~requests in $\sigma_{\set{u,v}}[t',t]$,
	and the moving cost,
	incurred by Alg is $\lambda + 2\alpha$.
  %That is, the cost of all requests in $\sigma_{\set{u,v}}[t',t]$.

@@

  \begin{enumerate}
  \item Case i)
	OPT has $\set{u,v}$ separated (in OPT's configuration) during $[t',t]$
	and therefore it pays $\lambda$.
	We charge OPT $\lambda/2 = \alpha$.
%	The remaining cost will be used in a latter case.

  \item Case ii)  
	OPT collocates $\set{u,v}$ some time during $[t',t]$.
	Any cost that we charge to OPT upon $\sigma_{t''}$,
	accounts for OPT's actions before $t''$.
	Therefore in this case,
	we charge OPT the (unaccounted) cost $\alpha$.

  \item Case iii)
	$t'' \geq 0$ and OPT has $\set{u,v}$ collocated before $t''$ and keeps them collocated during $[t'',t]$.
	In this case OPT doesn't pays any cost for requests in $\sigma_{\set{u,v}}[t'',t]$ and
	we can't charge it directly.
	However,
	OPT must pay $\lambda$ for all requests in $\sigma_{\set{u,v'}}[t'',t']$,
	 i.e.,
	 Case (i) applies to $\set{u,v'}$  at $t'$ and
 	we charge only $\lambda/2$ to OPT (at $t'$).
	This means,
	we can charge OPT an additional $\alpha$,
	i.e.~the remaining cost of requests $\sigma_{\set{u,v'}}[t'',t']$.
%	Intuitively, at $t$,  we charge $\alpha$ to OPT for separating $\set{u,v}$ (at $t'$) by Alg.
  \end{enumerate}

	It can be easily verified that any deviation from these cases only increases the cost to OPT.
	Thus,
	OPT pays at least $\alpha$ for each collocation performed by Alg.
%	which gives the competitive ratio 
%	$(\lambda + 2\alpha)/ \alpha = 4$.

	It remains to charge both Alg and OPT for requests that do not lead to collocations when the sequence ends.
	Consider  pairs $\set{u,v}$ and $\set{u,v'}$ and
	times $t'' < t' < t = t_m$ as defined previously.
	Assume $\sigma_{\set{u,v}}[t'', t_m] \leq \lambda - 1$.
	This means Alg has the pair separated at $t_m$
	and paid the (communication) cost $r_{\set{u,v}} := | \sigma_{\set{u,v}}[t'', t_m]|$.
	If OPT has the pair separated during $[t'',t_m]$ then it pays $r_{\set{u,v}}$.
	If OPT collocates the pair some time during $[t'', t]$ then analogous to Case (ii),
	we charge it $\alpha$.
	Otherwise,
	if OPT collocates the pair before $t''$,
	then we charge it $\alpha$ analogously to Case (iii).
	Hence, 
	$Cost(OPT)_{\set{u,v}} / Cost(Alg)_{\set{u,v}} =
	\max \set{ r_{\set{u,v}}/ \alpha \leq (\lambda-1) / \alpha , (\lambda + 2\alpha)/ \alpha} = 4$.
%	The claim follows from here.
\end{proof}

Missing:
\maciek{Reason that the cost sharing between cases C1 and C3 is possible. Unclear to me, why every time C3 happens there must be a corresponding C1. \mahmoud{done}}
\maciek{Missing the case of the first-time collocation. What if t' does not exist? What if t'' does not exist? 
\mahmoud{added remark for this case: $t''=0$}}

Minor:
\maciek{"It can be easily verified"... True, but avoid such formulation. Something along these lines: if opt migrates earlier than t', then additionally pays for X, if later than Y.}
\maciek{"Thus, OPT pays ... which gives the competitive ratio" It is too early to give the competitive ratio if in the next paragraph we consider yet another case. \mahmoud{done!}}

Tiny:
\maciek{$\sigma_t$ undefined (I know what it is, though). \mahmoud{added inline def. Do we have a global def for this?}}
\maciek{Define $C^t$ to use once \mahmoud{already defined! I don't get it.}}
\maciek{A word "Obviously" is unnecessary, and most people, including me are allergic to it in writing
\mahmoud{obviously that's true!}}
